### 2024-11-01 manuscript.qmd
Finish foreign aid donors section

---
title: "Donor Responses to Civil Society Repression"
subtitle: "Chapter for the *Handbook of Political Control*"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  orcid: 0000-0003-3506-7590
  url: https://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  orcid: 0000-0002-3948-3914
  url: https://www.andrewheiss.com
  email: aheiss@gsu.edu
  attributes:
    corresponding: true
bibliography: references.bib
toc: true
format:
  wordcount-html: 
    wordcount-banner: true
params:
  wordcount: |
    <strong>{{< words-sum body-note-ref >}} total words</strong>: {{< words-body >}} in the body • {{< words-ref >}} in the references • {{< words-note >}} in the notes
---

::: {.content-visible when-meta="wordcount-banner"}
{{< include _extensions/andrewheiss/wordcount/banner.html >}}
:::

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  echo = FALSE, include = FALSE, warning = FALSE, message = FALSE,
  fig.width = 6, fig.height = (6 * 0.618), fig.retina = 3,
  out.width = "100%", fig.align = "center"
)
```

```{r}
#| label: libraries-data
library(tidyverse)
library(tidygraph)
library(ggraph)
```

Given the exponential increase in administrative and legal repression of CSOs across the globe and its effectiveness at muzzling groups, how have donors responded? How do these responses differ across foreign government and private donors? How have local philanthropists and NGOs adapted and responded to such repression? What challenges do civil society groups continue to face despite these responses?

## Foreign aid donors

A rich literature on the purposes of foreign aid explores a broad collection of reasons why states transfer money and resources to others. Aid agencies and international organizations face competing interests when deciding how to allocate resources, which projects to undertake, and which countries to work with. Donor agencies rhetorically declare normative justifications for engaging in development, with recipient-focused development goals like poverty reduction, democratization, human rights promotion, and environmental protection [@HoefflerOutram:2011; @Kallman:2020]. In practice, though, aid is given for far more instrumental reasons based on donor countries' domestic and foreign policy–oriented strategic goals [@AlesinaDollar:2000; @Apodaca:2017; @Schuller:2012], including being responsive to popular democratic pressure [@HeinrichKobayashiLong:2018], insulating themselves from regional instability [@Bermeo:2017], expanding export markets [@BarthelNeumayerNunnenkamp:2014; @Mitchell:2002], influencing recipient states' domestic policies [@Apodaca:2013], and projecting donor country values [@MorgensternBrown:2022].

### Donor responses and conditionality

A key enforcement mechanism to ensure that recipient-country actions align with donor-country goals—and thus reduce the moral hazard inherent in donating resources—is to condition aid on the achievement of specific democratizing and economic goals [@Winters:2024; @CheesemanSwedlundOBrien-Udry:2024; @HoefflerJustino:2024; @Kilby:2024; @Koch:2015]. Many donor governments condition aid on human rights outcomes, using progress in respect for human rights as leverage to shape recipient-country behavior. Conditional aid sends dual signals: it serves as an instrumental "stick" to pressure governments to change policies and an expressive "flag" to signal donor commitment to human rights norms [@DasandiErez:2023]. However, this dual-purpose strategy often creates dilemmas—if aid is withdrawn, it may jeopardize humanitarian programs or worsen political conditions. Conversely, continuing aid might signal tolerance for rights violations, weakening the credibility of the donor’s human rights agenda.

The often-conflicting trends of both pursuing normative development goals and furthering foreign policy interests leads to mixed responses when recipient countries fail to adhere to aid conditions and violate human rights. Prior research finds conflicting relationships between human rights abuses and foreign aid, variously concluding that when confronted by recipient states that violate human rights, donor agencies punish repressive states by reducing, suspending, or withdrawing aid [@LebovicVoeten:2009], increasing aid (or withdrawing less aid) to repressive states as a way to counter state violence [@Demirel-PeggMoskowitz:2009; @Nielsen:2013; @HeinrichKobayashiLong:2018; @EsareyDeMeritt:2017], or make no measurable change in aid allocations [@Neumayer:2003; @Neumayer:2003b].

One reason for such mixed results is that the decision to withdraw or suspend aid in response to human rights abuses is often seen as a binary choice driven by self-interested motivations: a donor country that sees human rights abuses in a recipient state will consider its own foreign policy interests before deciding to either continue or curtail aid. In reality, however, the decision is far more complex and involves a wide range of possible motivations and responses. @CheesemanSwedlundOBrien-Udry:2024 offer a stylized decision tree of possible choices following recipient-country human rights violations—we provide a simplified version of their framework in @fig-agency-response-tree.

```{r}
#| label: fig-agency-response-tree
#| fig-cap: Possible donor- and recipient-country responses to human rights violations
#| include: true
#| fig-width: 6
#| fig-height: 3.8
#| out-width: 80%

nodes <- tribble(
  ~id, ~x, ~y,  ~step,
  1,   0,  2.5, "Violate human rights",
  2,   1,  4,   "Do nothing",
  3,   1,  3,   "Threaten",
  4,   1,  2,   "Withdraw\nor suspend",
  5,   1,  1,   "Shift aid",
  6,   2,  4,   "Do nothing",
  7,   2,  2.5, "Modify to improve\nrelationship",
  8,   2,  1,   "Pressure donors\nto back down"
)

edges <- tibble(
  from = c(1, 1, 1, 1, 2, 2, 2, 3, 3, 3, 4, 4, 4, 5, 5, 5),
  to = c(2, 3, 4, 5, 6, 7, 8, 6, 7, 8, 6, 7, 8, 6, 7, 8)
)

actors <- tribble(
  ~actor, ~label, ~x,
  "Recipient", "Recipient government\naction", 0,
  "Donor", "Donor government\nresponse", 1,
  "Recipient", "Recipient government\nresponse", 2
) |>
  mutate(
    xmin = x - 0.5,
    xmax = x + 0.5
  )

tidy_graph <- tbl_graph(nodes = nodes, edges = edges) |> 
  create_layout(layout = nodes)

ggraph(tidy_graph) +
  geom_rect(
    data = actors, aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = 4.2, fill = actor), inherit.aes = FALSE,
    alpha = 0.25
  ) +
  geom_rect(
    data = actors, aes(xmin = xmin, xmax = xmax, ymin = 4.2, ymax = Inf, fill = actor), inherit.aes = FALSE,
    alpha = 0.5
  ) +
  geom_text(
    data = actors, aes(x = x, y = 4.4, label = label), inherit.aes = FALSE,
    size = 3.5,
    fontface = "bold", lineheight = 1
  ) +
  geom_node_label(
    aes(x = x, y = y, label = step),
    label.r = unit(0, units = "pt"),
    size = 3.5, lineheight = 1, fontface = "bold"
  ) +
  geom_edge_link(
    aes(
      start_cap = label_rect(node1.step, padding = margin(5, 5, 5, 5, "pt")),
      end_cap = label_rect(node2.step, padding = margin(5, 5, 5, 5, "pt"))
    ),
    arrow = arrow(angle = 30, length = unit(6, "pt"), type = "closed"),
    edge_linewidth = 0.25, color = "grey40"
  ) +
  scale_x_continuous(expand = expansion(0)) +
  scale_fill_manual(values = c("grey40", "grey80"), guide = "none") +
  expand_limits(y = 0.9) +
  theme_void()
```

Donor countries have a range of possible responses to recipient repression, including doing nothing and ignoring the violation, threatening to change aid allocations, withdrawing or suspending aid, or shifting aid to other projects or programs. In turn, recipient governments can then respond by doing nothing and ignoring the threat or suspension, change their behavior and cease the repression to improve their relationship with the donor country, or pressure the donor country to back down and withdraw their complaint. In this simplified decision tree, there are 12 possible combinations of outcomes,^[@CheesemanSwedlundOBrien-Udry:2024 describe additional stages in their tree, with donors responding to recipients' responses, creating dozens of possible pathways.] and each pathway is conditional on a host of donor-country and recipient-country characteristics and contexts, including geopolitical dynamics in the recipient country's region, domestic political institutions within the donor country, public opinion in the donor country, coordination between coalitions of donors, the type of aid, and the type of human rights violation, among others [@CheesemanSwedlundOBrien-Udry:2024]. Such complexity makes it difficult to find consistent treatment effects in research that explores the overall effect of repression on aid—that is, we cannot definitively say that human rights causes an overall reduction or increase in aid. Instead, newer research looks at donor responses in more specific circumstances, exploring the factors that lead to specific response pathways [@Corwin:2023].

### Donor responses to civil society laws

Research on aid conditionality and agency responses to human rights abuses tends to look at more overt and violent forms of political oppression like respect for personal integrity, civil, and political rights [@Corwin:2023; @Neumayer:2003; @CheesemanSwedlundOBrien-Udry:2024], including arresting political activists, closing media outlets, suppressing protests, engaging in extrajudicial killings, and so on. The effects of more subtle administrative crackdown on aid, including the passage of anti-NGO laws and the constricting of civic space, are less explored and understood.

Civil society plays a crucial role in the distribution and implementation of international aid. Governmental aid agencies, such as the United States Agency for International Development (USAID) and the United Kingdom's Foreign, Commonwealth and Development Office (FCDO), channel substantial funding through NGOs in recipient countries. TODO: Something about the amount of money that goes through CS, perhaps highlighting just the US or UK. Larger international organizations also rely on domestic NGOs to carry out humanitarian and development projects. TODO: Something about the amount of money that the World Bank or UNHCR sends to NGOs. Outsourcing the provision of aid to NGOs allows donors to avoid aid capture in weak institutional environments and can lead to more effective program implementation [@Dietrich:2013; @CruzLabonneWright:2024]. Given the importance of NGOs in aid delivery, we have reason to expect that donor responses to civil society crackdown abroad are treated differently than more violent forms of repression, and thus likely to follow different pathways in @fig-agency-response-tree.

Donor countries have a range of possible responses when recipient countries engage in administrative crackdown. Recent research finds that national and multilateral aid agencies are responsive to anti-NGO laws, though often in inconsistent ways [@ChaudhryHeiss:2022a]. @ChristensenWeinstein:2013 show that aid from bilateral donors—or direct country-to-country aid—sees an average decrease of $25–\$50 million after recipient states pass laws that restrict the flow of foreign funds to NGOs. However, aid from international, multilateral agencies is not affected by these laws. Similarly @DupuyPrakash:2018 find that the adoption of a restrictive foreign funding law is associated with a 32% decline in aid from national, bilateral donors, but has no corresponding change in aid from international, multilateral donors.

One key reason for the divergence in the responses of bilateral and multilateral donors is the degree to which these donors rely on recipient-country NGOs to implement programs. Donor governments and agencies have different experiences with NGOs and correspondingly prioritize the openness of civil society in recipient countries in different ways. Multilateral donors like the World Bank and OECD tend to channel money directly to recipient governments rather than domestic NGOs and are thus more insulated from the effects of local civil society repression, which likely explains the lack of formal response to anti-NGO laws [@ChristensenWeinstein:2013; @DupuyPrakash:2018]. There is more variation, however, in how bilateral donors work with domestic NGOs. For example, @AllenFlynn:2018 find that the domestic political environments of donor countries shape how those countries channel their aid abroad. Because they tend to focus more on poverty alleviation initiatives, states with left-leaning governments typically channel more aid through recipient-country NGOs. In contrast, right-leaning governments are typically interested in promoting the economic and political interests of their own states, and accordingly prefer to channel aid directly through recipient-country governments [@AllenFlynn:2018]. Human rights abuses in recipient countries will shape the responses of donor countries depending on the saliency of the threat to aid provision: political repression that threatens business interests and investments will be more salient to right-leaning governments, while anti-NGO civil society administrative crackdown will likely have a stronger effect on left-leaning governments' responses, since domestic NGOs are central to donor countries' aid distribution strategies.

The nature of civil society restrictions also influences how official donor agencies respond. In their initial work on anti-NGO restrictions, @ChristensenWeinstein:2013 cataloged civil society laws into three broader categories—barriers to funding, barriers to entry, and barriers to advocacy—but they, and others like @DupuyPrakash:2018, dedicated the bulk of their analysis to looking at the effects of barriers to funding. This focus on funding restrictions is warranted, as most authoritarian civil society legal restrictions have been aimed specifically at reducing or controlling foreign funding [@Heiss:2019; @Carothers:2015a; @CarothersBrechenmacher:2014]. Moreover, the relationship between foreign aid and foreign funding restrictions appears to be bidirectional, as states are more likely to pass restrictive foreign funding laws in response to increases in bilateral aid [@DupuyRonPrakash:2016], especially if that aid arrives during competitive elections or times of political protest or instability [@Chaudhry:2022; @Heiss:2017]. Foreign funding restrictions and aid are thus linked: increasing aid leads to more funding restrictions [@DupuyRonPrakash:2016], which in turn lead to reductions in aid [@DupuyPrakash:2018].

Though the majority of anti-NGO laws have targeted foreign funding, other laws seek to impose barriers to entry, such as strict and burdensome registration requirements, and barriers to advocacy that explicitly limit NGO political participation [@Chaudhry:2022]. Donor countries respond differently to these types of oppression. @ChaudhryHeiss:2024 (see also @ChaudhryHeiss:2022a) find that bilateral donors are quite responsive to barriers to advocacy—on average, total foreign aid decreases by 30–40% (or \$100 in a typical country) following the passage of an advocacy-focused NGO law. @RightSpringmanWibbels:2024 find similar and more nuanced trends in donor responses: donors committed to political advocacy   (i.e. those that fund democracy and civil society promotion activities) reduce funding for advocacy programs by more than 70% in response to new restrictions, and the reduction in aid persists for several years.

Overall bilateral aid does not decrease significantly in response to new barriers to entry, but the distribution channels of that aid shift as donor countries channel more aid through domestic rather than foreign NGOs [@ChaudhryHeiss:2024]. USAID even explicitly encourages its program managers to rely more on domestic, in-country NGOs when facing a restrictive legal environment [@USAID:2014, 6–7]. In addition to changing the amounts and recipients of foreign aid, donor agencies often adjust the focus of the programs they fund, redirecting funding away from programs related to politically sensitive issues like human rights, anti-corruption, and democratic development and towards tamer topics that are more compatible with recipient-country preferences, such as education, health, and humanitarian services [@Bush:2015; @CheesemanDesrosiers:2023].

These shifts in funding patterns in response to restrictive civil society laws have had measurable effects on NGOs in recipient countries. Samantha Power, current administrator of USAID, has pledged to send 25% of USAID funding to local NGOs over the next few years, though practical obstacles remain. New local partners take longer to vet and train compared to established INGO partners. Turning to local organizations entails greater risk, as domestic NGOs can be easily shut down based on the whims of the government. Additionally,  communities often distrust local groups that accept Western aid, especially those focused on democracy assistance. Donor preferences have also changed. Over the past two decades, bilateral donors have increasingly provided short-term grants for projects that have easily measurable, quantifiable outputs [@Bush:2015]. This model may work well for civic groups working on relatively tame causes or exclusively development-focused goals, but more contentious organizations facing government crackdown may need flexible funding that does not bring immediate results or does not have easily measurable goals.

## Private donors and foundations

- Why individual donors are important
- Donor reactions are more than just binary, but there now fancy figure 1 decision tree yet, since (1) the existing literature is much smaller, and (2) individual donors have less power in relation to states. Countries don’t counter-react to donor shifts in behavior
- Different pathways exist, though, depending on donor preferences and organization characteristics. As with diverging studies in the effect of repression on formal aid, there are different effects. Experiment paper + simulation paper + conjoint paper (getting at donor and org features)

In other cases, donors shift their aid to different programs. - both good and bad. Bush and taming + USAID

While research in public policy examines philanthropic giving to NGOs, the discipline of political science and international relations tends to see philanthropy as less relevant. However, philanthropy can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. Understanding the dynamics of supporting international causes is important as philanthropy towards such causes can be harder to motivate because the number of recipients is larger and further removed from the donor.

While private sources can help raise such funds, official donor aid may not be easily replaceable by other sources. Testimony at a US Senate Committee on Foreign Relations hearing maintained that, “Neither remittances nor philanthropy can replace the expertise, scale, or agenda setting capacity of American foreign policy assistance.” Thus, despite promising new research on how to raise funds from private sources for besieged NGOs abroad, official aid donors also need to mount a coherent response to the shrinking civic space.