### 2024-11-05 README.qmd
Update readme

---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)

yaml_metadata <- rmarkdown::yaml_front_matter("manuscript.qmd")
title <- yaml_metadata$title
# abstract <- yaml_metadata$abstract
abstract <- "Given the increase in administrative and legal repression of civil society, how have their donors responded? How do these responses differ across foreign government and private donors? How have local philanthropists and NGOs adapted and responded to such repression? What challenges do civil society groups continue to face despite these responses?"
```

# `r title`

<!-- badges: start -->
[![Preprint](https://img.shields.io/badge/Preprint-10.31235%2Fosf.io%2F9gmhp-blue)](https://doi.org/10.31235/osf.io/9gmhp) [![OSF DOI](https://img.shields.io/badge/OSF-10.17605%2FOSF.IO%2FR97Y3-blue)](https://doi.org/10.17605/OSF.IO/R97Y3) [![Code DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.12817616.svg)](https://doi.org/10.5281/zenodo.12817616)
<!-- badges: end -->

[Andrew Heiss](https://www.andrewheiss.com/) • Andrew Young School of Policy Studies • Georgia State University  
[Suparna Chaudhry](https://www.suparnachaudhry.com/) • Lewis and Clark College

---

## Abstract

`r abstract`

## `r emoji_glue(":candy::bat:")`: Note on "butterscotch bat" project name

Because project titles change all the time with revisions, rewriting, and peer review, we used [{codename}](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames) internal-to-us project name that won’t change.

```{r generate-code-name, eval=FALSE}
library(codename)
codename_message()
#> code name generated by {codename} v.0.5.0. R version 4.4.1 (2024-06-14).

codename(seed = "handbook donors", type = "ubuntu")
#> [1] "butterscotch bat"
```

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

## Contributions and Code of Conduct

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.